---
title: "Final Project Markdown"
output: 
  html_document:
    code_folding: hide
---
Loading packages
```{r setup, include=FALSE}
library("dplyr")
library("tidyverse")
library("ggplot2")
library("hrbrthemes")
library("plotly")
library("viridis")
library("scales")
library("tigris")
library("tmap")
library("gghighlight")
library("hrbrthemes")
library("gsheet")
library("gghighlight")
```

## FTE Data

Upload data and create a dataframe called khn
```{r A, warning=F, message=F}
download.file("https://raw.githubusercontent.com/khnews/2020-underfunded-under-threat-data/master/data/05-local-health-departments-detail.csv", "05-local-health-departments-detail.csv")
khn <- rio::import("05-local-health-departments-detail.csv")
```

Create a new dataframe that filters just for NC
```{r}
nc <- khn %>%
  filter(state_code == "NC")
```

Select certain categories and filter out rows without fte data
```{r}
fte_data <- nc %>%
  select(`year`, `lhd_name`, `population`, `fte`, `fte_per_100000`, `population`) %>%
  filter(fte != "NA")
```

Find the average FTE per capita for the years we have available for each health department
```{r}
fte_data %>%
  group_by(lhd_name) %>%
  summarise(Avg_fte_per_100000 = mean(fte_per_100000)) %>%
  select(lhd_name, Avg_fte_per_100000) %>%
  arrange(desc(Avg_fte_per_100000))
```
Find the highest FTE per capita and FTE in 2017
```{r}
#FTE per capita
fte_data %>%
  filter(year == "2017") %>%
  arrange(desc(fte_per_100000))

#FTE
fte_data %>%
  filter(year == "2017") %>%
  arrange(fte)
```

Create a vector that lists counties in our sample size
```{r}
our_counties <- c("Alexander County Health Department", "Anson County Health Department", "Public Health Authority Cabarrus County", "Carteret County Health Department", "Catawba County Health Department", "Cleveland County Health Department", "Dare County Health Department", "Swain County Health Department", "Yadkin County Health Department", "Wayne County Health Department", "Granville-Vance District Health Department", "Greene County Health Department", "Montgomery County Health Department", "Pender County Health Department", "Person County Health Department", "Rowan County Health Department", "Harnett County Health Department", "Madison County Health Department", "Guilford County Department of Public Health", "Henderson County Health Department", "Iredell County Health Department", "Johnston County Health Department", "Lincoln County Health Department", "Pitt County Public Health Center", "Halifax County Department of Health", "Jones County Health Department", "Lenoir County Health Department", "Gaston County Health Department", "Hyde County Health Department", "Mecklenburg County Health Department",  "Rockingham County Health Department", "Chatham County Public Health Department", "Onslow County Health Department", "Franklin County Health Department", "Forsyth County Department of Public Health", "Orange County Health Department", "Brunswick County Health Department", "Richmond County Health Department", "Craven County Health Department", "Jackson County Department of Public Health", "Transylvania County Health Department", "Toe River District", "Union County Health Department", "Lee County Health Department", "Burke County Health Department", "Moore County Health Department", "Stanly County Health Department", "Surry County Health and Nutrition Center", "Wilkes County Health Department", "Martin Tyrell Washington District", "Wake County Human Services")
```

Find the average percent change in fte from 2007-2017
```{r}
fte_data %>%
  filter(lhd_name %in% our_counties) %>%
  filter(year %in% c('2007', '2017')) %>%
  pivot_wider(id_cols = -c(fte_per_100000, population), names_from = year, values_from = fte) %>%
  mutate(pct_change = 100 * ((`2017` - `2007`)) / `2007`) %>%
  mutate(avg = mean(`pct_change`)) %>%
  arrange(desc(pct_change))
```
## Expenditure Data Accounting for Inflation

Loading in necessary datasets.
```{r}
CompiledLHDExpenditures <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Zc10pam92Y1218F90eXn9ri7-a4GQI1vhzING33nNSI/edit#gid=0")
state_public_health_agencies <- read_csv("https://raw.githubusercontent.com/khnews/2020-underfunded-under-threat-data/master/data/01-state-public-health-agencies.csv")
```

Creating a dataset of inflation rates, no data for 2013.
```{r}
infl_rates <- 
  
  state_public_health_agencies %>%
  group_by(year) %>%
  summarize(total_raw = max(expenditures, na.rm=TRUE), total_infl = max(expenditures_infl, na.rm=TRUE)) %>%
  mutate(multiplier2019 = total_infl/ total_raw)
```

Using the infl_rates data to create a new dataframe of all LHD expenditures, adjusted for inflation.
```{r}
infl_CompiledLHDExpenditures <- 
  left_join(CompiledLHDExpenditures, infl_rates, by=c("year"="year")) %>%
  mutate(expenditures_infl = expenditures * multiplier2019, 
         expenditures_infl_per_capita = expenditures_infl / population) %>%
  dplyr::select(year, lhd_name, lhd_area_type, county_name, population, expenditures, expenditures_infl, expenditures_per_capita = per_capita_spending, expenditures_infl_per_capita)
```

Turning the years from a character string to an integer.
```{r}
infl_CompiledLHDExpenditures$year <- as.integer(infl_CompiledLHDExpenditures$year)
```

Pivoting the dataset so that the counties are in one row, and are separated by year columns.
```{r}
infl_nc_county_by_year <- infl_CompiledLHDExpenditures %>%
  filter(year >=2010, year<=2018) %>%
  pivot_wider(id_cols = county_name, names_from = year, values_from = expenditures_infl_per_capita, names_prefix = "year") %>%
  arrange(county_name)
```

Creating a line graph of every counties' per capita expenditures, adjusted for inflation.
```{r}
ggplotly(
  ggplot(filter(infl_CompiledLHDExpenditures,year>=2010, year<=2019),
         aes(x=year, 
             y=expenditures_infl_per_capita, 
             group=lhd_name, 
             color=lhd_name)) +
    geom_line() +
    scale_y_continuous()+
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Trends in Health Department Expenditures") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    )
)
```

*Overall trend is a slight decline, or at least the majority are staying the same.*

*From that chart we can tell that Franklin, Hyde, Dare, and the Martin-Tyrell-Washington District spend far more per person than any others.*

Comparing county expenditures to those of the North Carolina Department of Health and Human Services.
```{r}
ggplotly(
  ggplot(filter(infl_CompiledLHDExpenditures,year>=2010, year<=2019),
         aes(x=year, 
             y= expenditures, 
             group=lhd_name, 
             color=lhd_name)) +
    geom_line() +
    scale_y_continuous()+
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Trends in Health Department Expenditures") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    )
)
```

Loading in more necessary datasets.
```{r}
state_public_health_agencies <- read_csv("https://raw.githubusercontent.com/khnews/2020-underfunded-under-threat-data/master/data/01-state-public-health-agencies.csv")
state_local_government_spending <- read_csv("https://raw.githubusercontent.com/khnews/2020-underfunded-under-threat-data/master/data/02-state-local-government-spending.csv")
county_local_government_spending <- read_csv("https://raw.githubusercontent.com/khnews/2020-underfunded-under-threat-data/master/data/03-county-local-government-spending.csv")
local_health_departments_total_by_state <- read_csv("https://raw.githubusercontent.com/khnews/2020-underfunded-under-threat-data/master/data/04-local-health-departments-total-by-state.csv")
local_health_departments_detail <- read_csv("https://raw.githubusercontent.com/khnews/2020-underfunded-under-threat-data/master/data/05-local-health-departments-detail.csv", 
    col_types = cols(fips_place = col_character(), 
        fips_counties = col_character(), 
        lhd_note = col_character()))
```

Proving the claim that spending has dropped by 16% since 2010 across the U.S.
```{r}
#expenditures_overall <- 
  state_public_health_agencies %>% filter(!(state_code %in% c("MO", "MI", "TX", "WY"))) %>%
   filter(year %in% c(2010, 2018)) %>%
                group_by(year) %>%
                summarize(expenditures_infl = sum(expenditures_infl),
                                                                                                population = sum(population)) %>%
               
                mutate(expenditures_infl_per_capita = expenditures_infl/population)
```

Checking to see how much spending dropped in NC in the same 8 year span.
```{r}
state_public_health_agencies %>% 
  filter(state_code =="NC", year %in% c(2010, 2018)) %>%
                dplyr::select(year, expenditures_infl_per_capita)
```

*While expenditures per capita declined 16% in the United States, they fell 27% in North Carolina.*

*For every dollar spent on state agency public health nationally in 2010, NC spent 90 cents.*

*By 2018, that was only 77 cents per dollar.*

Rearranging the national dataframe from least to greatest, to see where NC ranks in expenditures per capita in 2018.
```{r}
  state_public_health_agencies %>% filter(!(state_code %in% c("MO", "MI", "TX", "WY"))) %>%
   filter(year == 2018) %>%
                dplyr::select(state_code, expenditures_infl_per_capita) %>% arrange(expenditures_infl_per_capita)
```
*At the state level in 2018, North Carolina spends less per person than all but 13 states. It spends about $72 per person.*

*While the spending at the state level has decreased, that has left counties to pick up the tab.*

Checking to see how much spending has dropped on average for NC counties.
```{r}
infl_CompiledLHDExpenditures %>% 
  filter(year %in% c(2010, 2018)) %>%
                group_by(year) %>%
                summarize(cum_expenditures = sum(expenditures_infl),
                                                                                                cum_population = sum(population)) %>%
                
                mutate(cum_expenditures_per_capita = cum_expenditures/cum_population)
```
Showing where each individual LHD ranks by pct_change in per capita.
```{r}
infl_nc_county_by_year %>% 
  filter(!is.na(year2010)) %>% 
  mutate(pct_change = (year2018 - year2010) / year2010) %>%
  arrange(pct_change) 
```
*The county with the biggest decline was Wake. From 2011 to 2014, it's per capita spending dropped nearly 60 percent.*

*On the other hand, Mecklenburg saw the biggest per capita increase -- 40% over those 8 years, and a jump of 28% from 2015 to 2016 alone.*

*Did that happen everywhere in 2015-2016?*

Looking at percent change numbers for all counties from 2015-2016. Mecklenburg had almost 14% more of an increase than the next highest, which was Richmond county at 14%.
```{r}
infl_nc_county_by_year %>% 
  dplyr::select(county_name, year2015, year2016) %>% 
  mutate(pct_change = (year2016 - year2015) / year2015) %>%
  arrange(pct_change) 
```
*Right next door to Wake, in more rural Johnston County in the same year, you had a 4% increase.*

*Mecklenburg dipped, too, but the rocketed back to see a 40% increase in per capita spending.*

Ordering the counties by how they ranked for their per capita expenditures in 2010.
```{r}
infl_nc_county_by_year %>% 
  filter(!is.na(year2010)) %>% 
  mutate(pct_change = (year2018 - year2010) / year2010) %>%
  arrange(year2010) 
```
*In 2010, Mecklenburg had one of the lowest per capita rates in the state -- similar to tiny Stanly and Burke counties. (Show an animated map change from 2010 to 2018 per capita.)*

*In 2010, Meck and Union were both near the bottom, spending similar amounts per person. But by 2018 as Meck continued to spend, Union' expenditures dropped nearly 32%.*

Sorting counties by their per capita expenditures in 2018.
```{r}
infl_nc_county_by_year %>% 
  filter(!is.na(year2010)) %>% 
  mutate(pct_change = (year2018 - year2010) / year2010) %>%
  arrange(year2018) 
```
*Creating a new dataframe that shows the change in population for each county.*
```{r}
pop_change <- 
  
  infl_CompiledLHDExpenditures %>%
  filter(year %in% c(2010, 2018)) %>%
  pivot_wider(id_cols = county_name, names_from = year, values_from = population, names_prefix = "year") %>%
  mutate(pop_diff = year2018-year2010, pct_diff = (year2018-year2010) / year2010)  %>%
  mutate(change = case_when(pop_diff > 0 ~ "increase", 
                            pop_diff <=0 ~ "decrease")) %>%
 arrange(county_name)
pop_decrease <- 
  pop_change %>%
  filter(change == "decrease")
```

*Some counties saw their per capita spending increase because their population plummeted even while state population grew about 10%.*

*For example, Richmond County's per capita spending increased nearly 15% even while it's population decreased 3%.*

Creating a new dataframe that shows counties that increased in per capita spending, and then finding where they intersect with counties that decreased in population. The only two counties where both conditions are true are Cleveland and Richmond counties.
```{r}
increase_per_cap<- 
  
  infl_nc_county_by_year %>% 
  filter(!is.na(year2010)) %>% 
  mutate(pct_change = (year2018 - year2010) / year2010) %>%
  mutate(change = case_when(pct_change > 0 ~ "increase", 
                            pct_change < 0 ~ "decrease")) %>%
  filter(change == "increase") %>% arrange(county_name) %>%
  dplyr::select(county_name, pct_change, change) %>%
  arrange(county_name)
intersect(increase_per_cap$county_name, pop_decrease$county_name)
```
Putting what was found above in a more readable chart, along with the corresponding numbers.
```{r}
inner_join(increase_per_cap, pop_decrease, by=c("county_name"="county_name")) %>%
  dplyr::select(county_name, per_cap_spending_increase = pct_change, pop_pct_decrease = pct_diff)
```
*Which counties saw decreased spending?*

Creating a dataframe showing exclusively counties that had an overall decrease in expenditures from 2010-2018.
```{r}
raw_expenditure_change <- 
  infl_CompiledLHDExpenditures %>%
  filter(year %in% c(2010, 2018)) %>%
  pivot_wider(id_cols = county_name, names_from = year, values_from = expenditures, names_prefix = "year") %>%
  mutate(expend_diff = year2018-year2010, raw_diff = year2018-year2010) %>%
  mutate(change = case_when(expend_diff > 0 ~ "increase", 
                            expend_diff <=0 ~ "decrease")) %>%
  filter(change == "decrease") %>% arrange(county_name)

raw_expenditure_change
```
Did any counties see decrease in spending but increase in per-capita?
```{r}
intersect(increase_per_cap$county_name, raw_expenditure_change$county_name)
```
*No counties had a decrease in spending while also having an increase in per-capita spending.*

*When looking at Full-Time-Employees (FTEs), the year range has to change from 2007-2017, because the state does not report FTE data in even-numbered years, and the most recent report was in 2017.*

*Because the FTE is reported, we do have data for all 85 North Carolina local health departments.*

Creating a dataframe to better examine the changes in FTEs from 2009-2017.
```{r}
nc_fte_change <- 
  local_health_departments_detail %>%
  filter(state_code == "NC", year %in% c(2009, 2017))%>%
  pivot_wider(id_cols = c(lhd_name, county_name), names_prefix = "year", names_from = year, values_from = fte) %>%
  mutate(change = year2017 - year2009, pct_change = (year2017 - year2009)/ year2009 ) %>%
  arrange(desc(pct_change))
```

*Of the 85 LHDs, 62 had fewer people working at them in 2017 than in 2009.*

*Wake County has cut 268 FTEs over that decade -- 30% of its workforce. Meanwhile, Mecklenburg County more than doubled its workforce from 379 in 2007 to 825 in 2017. That's the biggest increase in terms of both raw numbers of new employees as well as percent increase.*

Ranking each LHD by their pct_change in FTE.
```{r}
local_health_departments_detail %>%
  filter(state_code == "NC", year %in% c(2009, 2017))%>%
  pivot_wider(id_cols = lhd_name, names_prefix = "year", names_from = year, values_from = fte) %>%
  mutate(change = year2017 - year2009, pct_change = (year2017 - year2009)/ year2009 ) %>%
  arrange(desc(pct_change)) %>%
  filter(pct_change < -0.33)
```
*Wake's cuts were biggest in terms of raw numbers, but 13 health departments across the state saw their staffing drop by at least a third. Hyde County lost two thirds of its FTE. The Toe River Health District went from 200 to 60 people.*

Filtering each county so only increases in FTEs are shown.
```{r}
  infl_CompiledLHDExpenditures %>%
  filter(year %in% c(2009, 2017)) %>%
  pivot_wider(id_cols = county_name, names_from = year, values_from = population, names_prefix = "year") %>%
  mutate(pop_diff = year2017-year2009, pct_diff = (year2017-year2009) / year2009)  %>%
  mutate(change = case_when(pop_diff > 0 ~ "increase", 
                            pop_diff <=0 ~ "decrease")) %>%
  filter(change == "increase") %>% arrange(county_name)
```
Filtering to find all the health departments that had their number of FTEs decrease.
```{r}
nc_fte_change %>% filter(change <0 ) %>% arrange(lhd_name)
```
*Brunswick County's population exploded by more than a third over those 10 years. At the same time, it's public health staff was cut by 41%.*

*23 other counties or districts that gained population but lost public health FTEs:*
* Avery, Mitchell, Yancey (Toe River) 
* Catawba 
* (Public Health Authority of) Cabarrus 
* Craven 
* Dare 
* Forsyth
* Franklin
* Granville & Vance
* Guilford
* Iredell
* Jackson
* Lee
* Lincoln
* Moore
* Person
* Rowan
* Swain
* Transylvania
* Union
* Wayne

Ranking all the LHD's by their number of FTEs in 2017.
```{r}
  local_health_departments_detail %>%
  filter(state_code == "NC", year %in% c(2017)) %>%
  dplyr::select(lhd_name, fte) %>%
  arrange(fte)
```
Ranking the counties by their per capita expenditures.
```{r}
infl_CompiledLHDExpenditures %>%
  filter(year >= 2018) %>%
  dplyr::select(county_name, year, expenditures_infl_per_capita) %>%
  arrange(expenditures_infl_per_capita)
```
*This isn't just about poverty. Union County has the  highest median household income in the state -- about $86,000 -- but in 2018 spent less per person on public health than any other county -- just $32.92.*

Creating a plot that shows the expenditures per capita for each county, with the size of the points being equal to the county population in 2018.
```{r}
oneyear <- infl_CompiledLHDExpenditures %>% 
filter(year==2018)
library(ggrepel)
average_pcs<- oneyear%>% summarize(mean(expenditures_infl_per_capita))
ggplot(oneyear,
       aes(x=1, y = expenditures_infl_per_capita)) +
  geom_point(aes(size=population)) +
  theme_bw() +
    xlab("") +
    geom_text_repel(aes(label=county_name)) +
  geom_hline(yintercept=	
83.05706, color="red")
```

Taking a closer look at Jones County's FTE data. Jones is near the bottom in FTE ranking, and multiple interviews have noted Jones as a good example of an underfunded and understaffed health department.
```{r}
local_health_departments_detail %>% filter(state_code == "NC", lhd_name =="Jones County Health Department") %>%
  dplyr::select(year, fte) %>%
  arrange(year)
```
Getting the inflation rates for each year.
```{r}
#infl_rates <- 
  state_public_health_agencies %>%
  group_by(year) %>%
  summarize(total_raw = max(expenditures, na.rm=TRUE), total_infl = max(expenditures_infl, na.rm=TRUE)) %>%
  mutate(multiplier2019 = total_infl/ total_raw)
```

There was no state expenditure data collected in 2012, at least in this dataset.
```{r}
state_public_health_agencies %>%
select(year, expenditures, expenditures_infl) %>%
filter(year == 2012) %>%
arrange(year)
```

Adding the inflation rates to the inflation-adjusted CompiledLHDExpenditures dataframe.
```{r}
#infl_infl_CompiledLHDExpenditures <- 
left_join(infl_CompiledLHDExpenditures, infl_rates, by=c("year"="year")) %>%
  mutate(expenditures_infl = expenditures * multiplier2019, 
         expenditures_infl_per_capita = expenditures_infl / population) %>%
  select(year, lhd_name, lhd_area_type, county_name, population, expenditures, expenditures_infl, expenditures_per_capita = expenditures_infl_per_capita, expenditures_infl_per_capita)
```

## Graphing

Loading data
```{r}
library(gsheet)
CompiledLHDExpenditures <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Zc10pam92Y1218F90eXn9ri7-a4GQI1vhzING33nNSI/edit#gid=0")
```

### Lollipop chart showing counties that have increased in per capita spending vs. decreased from 2010-2018

Loading the data
```{r}
infl_nc_county_by_year <- infl_CompiledLHDExpenditures %>%
  filter(year >=2010, year<=2018) %>%
  pivot_wider(id_cols = county_name, names_from = year, values_from = expenditures_infl_per_capita, names_prefix = "year") %>%
  arrange(county_name)
```

Create a new dataframe with a new column that shows the percent change in spending from 2010-2018
```{r}
PerCapitaSpendingChange <- infl_nc_county_by_year %>% 
  filter(!is.na(year2010)) %>% 
  mutate(pct_change = (year2018 - year2010) / year2010) 
```

Create a lollipop chart that visualizes each health department's percent change in expenditures
```{r}
ggplot(PerCapitaSpendingChange,
       aes(x=pct_change, y=fct_reorder(county_name, pct_change))) +
  geom_segment(
    aes(x=0,
        y=fct_reorder(county_name,pct_change),
        xend= pct_change,
        yend=fct_reorder(county_name, pct_change)),
    color="gray50")+
  geom_point(color="#1d91c0")+
  labs(x="Percent Change in Per Capita Spending", y="County Health Department", 
       title = "Percent Change in Per Capita Spending from 2010-2018",
       caption = "Data Source: County Finance Records (Did not receive data from 34 Health Departments)") +
  theme_minimal()+
   theme(panel.border = element_blank(),
        panel.grid.minor = element_blank()
   )
```

Save the image as a png file
```{r}
ggsave("per-capita-spending-lollipop-chart.png")
```

### Lollipop chart of FTE change per health department from 2009-2017

Loading the data
```{r}
library(gsheet)
FTE_change_revised_names_start2009 <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1-uDrdN2JbvSeepLtUnSRROnJ7ci9rJkxq9HC8iTvDaA/edit#gid=0")
```

Creating the chart
```{r}
ggplot(FTE_change_revised_names_start2009,
       aes(x=pct_change, y=fct_reorder(lhd_name, pct_change))) +
  geom_segment(
    aes(x=0,
        y=fct_reorder(lhd_name,pct_change),
        xend= pct_change,
        yend=fct_reorder(lhd_name, pct_change)),
    color="gray50")+
  geom_point(color="#1d91c0")+
  labs(x="Percent Change in FTEs", y="County Health Department", 
       title = "Percent Change in Number of FTEs from 2009-2017",
       caption = "Data Source: KHN FTE Dataset") +
  theme_minimal()+
   theme(panel.border = element_blank(),
        panel.grid.minor = element_blank()
   )
```

Save the image as a png file
```{r}
ggsave("FTE-change-lollipop-chart-2009.png", width= 7, height = 10)
```

### Change in per capita spending for Mecklenburg, Wake, Union and Johnston 

Creating a new dataframe
```{r}
MeckWakeUnionJohnston <- CompiledLHDExpenditures %>%
  filter(county_name %in% c("Mecklenburg", "Wake", "Union", "Johnston"), year >=2010, year<=2018)
```

Creating a line chart
```{r}
ggplotly(
ggplot(MeckWakeUnionJohnston) +
  geom_line(mapping = aes(x = year, y = per_capita_spending, color = county_name)) +
  labs(x = "Year", y = "Per Capita Spending", title = "Changes in Health Department Spending per Capita from 2010-2018", color = "County", caption = "Data Source: County Finance Records") +
 scale_y_continuous(labels= scales::dollar_format())+
  theme_minimal() 
)
```
### Line chart showing change in health department spending per capita from 2014-2018

Creating a new dataframe
```{r}
infl_MeckWakeUnionJohnston <- infl_CompiledLHDExpenditures %>%
  filter(county_name %in% c("Mecklenburg", "Wake", "Union", "Johnston"), year >=2014, year<=2018)
```

Creating a line chart
```{r}
ggplotly(
ggplot(infl_MeckWakeUnionJohnston) +
  geom_line(mapping = aes(x = year, y = expenditures_infl_per_capita, color = county_name)) +
  labs(x = "Year", y = "Per Capita Spending", title = "Changes in Health Department Spending per Capita from 2014-2018", color = "County", caption = "Data Source: County Finance Records, adjusted for inflation") +
 scale_y_continuous(labels= scales::dollar_format())+
  theme_minimal() 
)
```

### Creating an interactive map

Colored in by per capita spending
On hover also shows population, expenditure and fte percent change

Changing the dataframe to calculate percent change
```{r}
change_infl_CompiledLHDExpenditures <- infl_CompiledLHDExpenditures %>%
  filter(year %in% c(2010, 2018)) %>%
  dplyr::select(year,
         county_name,
         population,
         expenditures,
         expenditures_infl,
         expenditures_per_capita,
         expenditures_infl_per_capita) %>%
    pivot_wider(
      id_cols = county_name, 
                names_from = year, 
                values_from =c(population, expenditures,
         expenditures_infl,
         expenditures_per_capita,
         expenditures_infl_per_capita)
    )

#Calculating percent change
change_infl_CompiledLHDExpenditures<- change_infl_CompiledLHDExpenditures %>%
  mutate(
    pct_change_expenditures_infl = (expenditures_infl_2018 - expenditures_infl_2010)/ expenditures_infl_2010,
    pct_population_change = (population_2018 - population_2010) / population_2010,
    pct_change_expenditures_capita_infl = (expenditures_infl_per_capita_2018 - expenditures_infl_per_capita_2010) / expenditures_infl_per_capita_2010
  )


#Join the two dataframes
nc_fte_change<- nc_fte_change%>% mutate(only_county_name = str_sub(county_name,end = -8))
  
a_MappingData<-  left_join(nc_fte_change, change_infl_CompiledLHDExpenditures, by=c("only_county_name"="county_name"))


#Clean up some column names and reorder them
a_MappingData<- a_MappingData %>%
  rename (fte_pct_change = pct_change
          ) %>%
  dplyr::select(lhd_name,
         county_name,
         only_county_name,
         fte_pct_change,
         pct_change_expenditures_infl,
         pct_population_change,
         everything())


```

Added rows for multi-county data in Google Sheets so those counties get filled in on map. Loading in that data.
```{r}
library(gsheet)
MultiCountyData <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Az71mA2PJJ7oJskn_A3eAdfyY5loHcT7CT7g2523odE/edit#gid=0")
```

Downloading shapefile of NC counties
```{r}
library(tigris)
options(tigris_use_cache = TRUE)
nc_counties <- counties("NC")
```

Joining data with fte data
```{r}
test_pct_capita_map <- left_join(nc_counties, MultiCountyData, by=c("NAME"="only_county_name"))
```

Leaflet map where health districts get filled in
```{r}
library(leaflet)
library(scales)
pal <- colorNumeric("YlGnBu", domain=test_pct_capita_map$pct_change)

#Setting up the pop up text
popup_pctchange <- paste0(
  "<strong>", test_pct_capita_map$NAMELSAD,
  "</strong><br /> Population Change: ", as.character(
    percent(
      round(test_pct_capita_map$pct_population_change, digits=1),
      scale=100,
      big.mark = ",") ),
  "<br /> Expenditure Change: ",
  as.character(
    percent(
      round(test_pct_capita_map$pct_change_expenditures_infl, digits=1),
      scale=100,
      big.mark = ",") ),
    "<br /> FTE Change: ",
  as.character(
    percent(
      round(test_pct_capita_map$fte_pct_change, digits=1),
      scale=100,
      big.mark = ",") )
)


leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng=-79.177556, lat=35.481333, zoom = 6) %>%
    addPolygons(data = test_pct_capita_map, 
              fillColor = ~pal(test_pct_capita_map$pct_change_expenditures_capita_infl), 
              fillOpacity = 0.9, 
              weight = 0.2, 
              smoothFactor = 0.2, 
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE), 
              popup = ~ popup_pctchange) %>%
  addLegend(pal = pal,
            opacity = 0.9,
            values = test_pct_capita_map$pct_change_expenditures_capita_infl * 100,
            position = "bottomleft",
            title = "$ per capita", 
            labFormat = labelFormat(suffix = "%"),
            na.label = "No Data") 

```

### Line chart showing change in national per capita spending at state level and NC per capita spending at state level

```{r}
library(tidyverse)
state_per_capita <- read_csv("01-state-public-health-agencies.csv")
```

```{r}
ggplot(state_per_capita, aes(x=year, y=expenditures_infl_per_capita, group = state_code, color = state_code)) +
  geom_line() + 
  scale_fill_viridis(discrete = TRUE) +
  scale_x_continuous(breaks=c(2014, 2015, 2016, 2017, 2018, 2019)) +
  scale_y_continuous(labels = dollar) +
  labs(title = "Per Capita Spending by State Public Health Agencies, 2014-2018", y = "Per Capita Spending", x = "Year", caption="Note: No data for 2011-2014, Source: https://github.com/khnews/2020-underfunded-under-threat-data/blob/master/data/01-state-public-health-agencies.csv") +
  theme_ipsum() +
  gghighlight(state_code == "NC") +
 coord_cartesian(xlim = c(2014, 2018), ylim = c(0, 200))
```

