---
title: "graphs"
author: "Rachel Crumpler"
date: "10/25/2020"
output: html_document
---
Loading data
```{r}
library(gsheet)
CompiledLHDExpenditures <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Zc10pam92Y1218F90eXn9ri7-a4GQI1vhzING33nNSI/edit#gid=0")
```

Loading packages
```{r}
library("dplyr")
library("tidyverse")
library("ggplot2")
library("hrbrthemes")
library("plotly")
library("viridis")
```


```{r}
pacman::p_load('dplyr', 'tidyr', 'gapminder',
               'ggplot2',  'ggalt',
               'forcats', 'R.utils', 'png', 
               'grid', 'ggpubr', 'scales')
```


Filtering Jones data
```{r}
Jones <- CompiledLHDExpenditures  %>%
  filter(county_name =="Jones")
```

#Interactive Jones County Line Chart

**Need to refine them by decreasing title font size and centering it
```{r}
ggplotly(
  ggplot(Jones,
    aes(x=year, 
        y=expenditures)) +
    geom_line(color="#69b3a2") +
    scale_y_continuous(labels= scales::dollar_format()) +
    labs(title="Trend of Jones County Health Department Spending",
         caption= "Data Source: Jones County Annual Expenditure Records") +
    ylab("Annual Expenditures") +
    xlab("Year") +
    theme_ipsum()
)
```



```{r}
ggsave("Jones-line-chart.png")
```



****THIS WONT SAVE FOR SOME REASON --I think it has something to do with graph coming from an object
Interactive Jones line chart
```{r}
Jones_graph <- plot_ly(Jones, x = ~year, y = ~expenditures, type = 'scatter', mode = 'lines')

layout(Jones_graph , title= "Trend of Jones County Health Department Spending" , xaxis = list(title= "Year"), yaxis = list(title= "Expenditures ($)"), caption="Data Source: Jones County Annual Expenditure Records")

```

#Interactive Jones County area chart
```{r}
  ggplotly(
  ggplot(Jones,
         aes(x=year, 
             y=expenditures)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    scale_y_continuous(labels= scales::dollar_format())+
    labs(title="Trend of Jones County Health Department Spending", caption= "Data Source: Jones County Annual Expenditure Records", y="Annual Expenditures", x="Year") +
    theme_ipsum()
  )
```

```{r}
ggsave("Jones-area-chart.png")
```


Filtering Mecklenburg data
```{r}
Mecklengburg <- CompiledLHDExpenditures  %>%
  filter(county_name =="Mecklenburg")
```

#Mecklenburg County Line Graph
```{r}
ggplotly(
  ggplot(Mecklengburg,
    aes(x=year, y=expenditures)) +
    geom_line(color="#69b3a2") +
    scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
    labs(title="Trend of Mecklenburg County Health Department Spending", caption= "Data Source: Jones County Annual Expenditure Records", y="Annual Expenditures", x="Year") +
    theme_ipsum()
)
```

```{r}
ggsave("Mecklenburg-line-graph.png")
```


Graphs shows trends of all health departments

```{r}
ggplotly(
  ggplot(CompiledLHDExpenditures,
         aes(x=year, 
             y=expenditures, 
             group=lhd_name, 
             color=lhd_name)) +
    geom_line() +
    scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Trends in Health Department Expenditures") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    )
)
```

Legend blocks the graph
```{r}
CompiledLHDExpenditures %>%
  ggplot( aes(x=year, y=expenditures, group=lhd_name, color=lhd_name)) +
    geom_line() +
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Trends in LHD Expenditures") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    )
```

Interactive graph of health departments under 2 million

```{r opts_chunk$set(out.height='7500px', dpi=200)}
ggplotly(
  ggplot(CompiledLHDExpenditures, aes(
    x=year, 
    y=expenditures, 
    color=lhd_name)) +
    geom_line() +
    scale_y_continuous(labels= scales::comma, limits= c(0, 20000000))+
    scale_fill_viridis(discrete = TRUE) +
    ggtitle("Trends in LHD Expenditures") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    )
)
```
--------USE THIS GRAPH

Interactive line graph showing all counties
*Need to fix what shows up in pop up on hover
```{r}
ggplotly(
  ggplot(CompiledLHDExpenditures, aes(
    x=year, 
    y=expenditures, 
    color=lhd_name)) +
    geom_line(size=.1) +
    scale_fill_viridis(discrete = TRUE) +
    scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
    labs(title="Trends in Local Health Department Expenditures", y= "Annual Expenditures", x= "Year", caption="Data Source: County Health Department Finance Records") +
    theme_ipsum()+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    )
)
```
To view graph better, click on first button in top right corner with chart of window with error

```{r}
ggsave("LHD-expenditure-line-chart.png")
```


--------------------GGHIGHLIGHT-----------------------

```{r}

# This helps us see what years we have the most data.

CompiledLHDExpenditures %>% group_by(year) %>% count()

 

#for each lhd, what is the oldest and newest data we have

max_min_years<- CompiledLHDExpenditures %>% group_by(lhd_name) %>% summarize(oldest= paste0("year_",min(year)), newest=paste0("year_",max(year)))

 

# we’re going to need to create a column that calculates change over time

pivoted_CompiledLHDExpenditures <- CompiledLHDExpenditures %>%

arrange(year) %>%

  pivot_wider(id_cols = lhd_name, names_from = year, names_prefix="year_", values_from=per_capita_spending)

 

joined <- left_join(max_min_years, pivoted_CompiledLHDExpenditures) %>% as.data.frame()

```



```{r}

# here we calculate change over time (tried to do dynamic calculation based on max and min years for each lhd, but no luck…) and also create a column that just says whether the county was up or down. I decided to look just at per-capita spending to account for different sizes of counties.


pivoted_CompiledLHDExpenditures <- pivoted_CompiledLHDExpenditures %>% mutate(change=year_2019 - year_2009) %>%

  filter(!is.na(change)) %>%

  mutate(trend = case_when(change >0 ~ "up",

                           change < 0 ~ "down",

                           TRUE ~ "no change"

      ))

```

```{r}
#24 counties down, 18 up. Hmmm…. Might be interesting to map these…

pivoted_CompiledLHDExpenditures %>% group_by(trend) %>% count()
```

 
```{r}

#we had to pivot wider to calculate across the rows of years, but now we have to pivot longer so we can do the line charts the way we want….

 

unpivoted <- pivoted_CompiledLHDExpenditures %>% pivot_longer(cols = starts_with("year_"), names_prefix="year_", names_to="year", values_to="per_capita_spending")

```

 

#Per Capita Spending Line Graph of all Health Departments
```{r}

# here we use gghighlight to show only the up counties.

#install.packages("gghighlight")

library(gghighlight)

ggplotly(

  ggplot(unpivoted, aes(

    x=year,

    y=per_capita_spending,

    color=lhd_name,

    group=lhd_name)) +

    geom_line() +

    scale_fill_viridis(discrete = TRUE) +

scale_y_continuous(labels = dollar) +
  
  scale_x_discrete(breaks=c(2000, 2005, 2010, 2015, 2020))+

    labs(title="Local Health Departments with Increases in Per Capita Spending", y= "Per Capita Spending", x= "Year", caption="Data Source: County Health Department Finance Records") +
    theme_ipsum()+

gghighlight(trend =="up") +

    theme(

      legend.position="none",

      panel.spacing = unit(0.1, "lines"),

      strip.text.x = element_text(size = 8),

      plot.title = element_text(size=14)

    )

)


```

```{r}
ggsave("increases-per-capita-spending.png")
```

 
#Biggest decrease in per capita spending
```{r}

#22 is too many to look at, so let’s just focus on the counties that had, say, a decrease in per-capita spending of more than $25 per person…. And we also just show years 2009-2019…

 

plot<- unpivoted %>% 
  filter(year>2008 & year < 2020 & per_capita_spending >0) %>%

  ggplot(aes(

    x=year,

    y=per_capita_spending,

   color=lhd_name,

    group=lhd_name)) +

    geom_line() +

    scale_fill_viridis(discrete = TRUE) +

scale_y_continuous(labels = dollar) +

 labs(title="Biggest Declines in Per Capita Spending, 2009-2019", subtitle = "Counties decreased by more than $25", y="Per Capita Spending", x= "Year", caption="Data Source: County Health Department Finance Records") +

gghighlight(change < -25)+

theme_ipsum() 

 
ggplotly(plot)


```

```{r}
ggsave("biggest-declines-per-capita-spending.png")
```


#Biggest Increase in Per Capita Spending
```{r}
#here are the counties that increased the most…

plot<- unpivoted %>% 
  filter(year>2008 & year < 2020 & per_capita_spending > 0) %>%

 

  ggplot(aes(

    x=year,

    y=per_capita_spending,

    color=lhd_name,

    group=lhd_name)) +

    geom_line() +

    scale_fill_viridis(discrete = TRUE) +

scale_y_continuous(labels = dollar) +

  labs(title="Biggest Increases in Per Capita Spending, 2009-2019", subtitle = "Counties increased by more than $25", y="Per Capita Spending", x= "Year", caption="Data Source: County Health Department Finance Records") +

gghighlight(change > 25) +
  
theme_ipsum()

 

ggplotly(plot)

 
```
 
```{r}
ggsave("biggest-increases-per-capita-spending.png")
```
 
-----------------FACETED GRAPHS--------------------

#Making line graph of each individual LHD
```{r}
CompiledLHDExpenditures %>%
  ggplot(
    aes(x=year, 
        y=expenditures, 
        group=county_name, 
        fill=county_name)) +
    geom_line(color="#69b3a2") +
    scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
    scale_x_discrete(breaks=c(2004, 2008, 2012, 2016))+
    scale_fill_viridis(discrete = TRUE) +
    labs(title="Trends in Health Department Expenditures", 
         y="Annual Expenditures",
         x="Year", 
         caption="Data Source: County Finance Records") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    ) +
    facet_wrap(~lhd_name)
```
```{r}
ggsave("faceted-expenditure-graph.png")
```

#Isolating some of the LHDs for more specific line chart comparison (these are the counties we interviewed a health director)
**Need to fix legend title
```{r}
linegraph <- CompiledLHDExpenditures %>% 
  filter(county_name %in% c("Jones", "Granville & Vance", "Madison", "Mecklenburg"))

# Plot
linegraph %>%
  ggplot( aes(x=year, 
              y=expenditures, 
              group=lhd_name, 
              color=lhd_name)) +
    geom_line() +
    scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
    scale_color_viridis(discrete = TRUE) +
    theme(legend.position="bottom"
    ) +
    labs(title="Comparing Expenditures of Health Departments We Interviewed", y="Annual Expenditures") +
    theme_ipsum()

```
```{r}
ggsave("expenditures-interviewed-counties.png")
```

#This graph compares some of the LHD expenditures (can be adjusted for when we find counties we want to put next to each other for closer comparison)
```{r}
linegraph <- CompiledLHDExpenditures %>% 
  filter(county_name %in% c("Swain","Jones", "Catawba", "Hyde", "Madison", "Mecklenburg"))

# Plot
linegraph %>%
  ggplot(aes(x=year, y=expenditures, group=lhd_name, color=lhd_name)) +
    geom_line() +
    scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
    scale_color_viridis(discrete = TRUE) +
    theme(
      legend.position="none",
      plot.title = element_text(size=14), 
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    ggtitle("Comparing LHD Expenditures") +
    theme_ipsum()+
    facet_wrap(~lhd_name)
```

Jones County fte per capita change over time
```{r}
Jones_graph_fte_pc <- plot_ly(jones, x = ~year, y = ~fte_per_100000, type = 'scatter', mode = 'lines')

Jones_graph_fte_pc <- Jones_graph_fte_pc %>%
  layout(title = "Jones County FTE Per Capita 1999-2017", xaxis = list(title = "Year"), yaxis = list(title = "FTE Per Capita"))

Jones_graph_fte_pc
```
<<<<<<< HEAD
Scatterplot of population and fte per capita over time, highlighting where fte per capita > 500 
```{r}
ggplot(fte_per_capita, aes(x=population, y=fte_per_100000)) + 
  geom_point(size= 1) +
   facet_wrap(~year, ncol=4) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  gghighlight(fte_per_100000 > 500) +
  labs(title = "FTE Per Capita and Population Over Time", x = "Population", y = "Employees Per Capita")

#NOT USING THIS
```
```{r}
ggsave("fte_pc_and_population_graph.png")
```

The percent change in full-time equivalent employees (FTE) from 2009-2017 
```{r}
nine_seventeen_fte_change <- fte_per_capita2 %>%
  filter(lhd_name %in% our_counties) %>%
  filter(year %in% c('2009', '2017')) %>%
  pivot_wider(id_cols = -c(fte_per_100000), names_from = year, values_from = fte) %>%
  mutate(pct_change = 100 * ((`2017` - `2009`)) / `2009`) %>%
  arrange(desc(pct_change))

library(forcats)
ggplot(nine_seventeen_fte_change, aes(x=pct_change, y=fct_reorder(lhd_name, pct_change))) +
geom_segment(aes(x = 0, xend = pct_change, yend = fct_reorder(lhd_name, pct_change)), color = "gray50") +
geom_point(color="#1d91c0")+
labs(title = "FTE Percent Change From 2009-2017", x ="Percent Change in FTE", y = "County Health Department", caption = "Data Source: Association of State and Territorial Health Officials")+
theme_minimal()+
theme(panel.border = element_blank(),
panel.grid.minor = element_blank()
)
```
```{r}
ggsave("fte_perc_change.png")
```
=======

---------------------FILLED LINE--------------------------

```{r}
oh_nine <- CompiledLHDExpenditures %>%
  filter(year == "2009")

ten <- CompiledLHDExpenditures %>%
  filter(year == "2010")

eleven <- CompiledLHDExpenditures %>%
  filter(year == "2011")

twelve <- CompiledLHDExpenditures %>%
  filter(year == "2012")

thirteen <- CompiledLHDExpenditures %>%
  filter(year == "2013")

fourteen <- CompiledLHDExpenditures %>%
  filter(year == "2014")

fifteen <-CompiledLHDExpenditures %>%
  filter(year == "2015")

sixteen <- CompiledLHDExpenditures %>%
  filter(year == "2016")

seventeen <- CompiledLHDExpenditures %>%
  filter(year == "2017")

eighteen <- CompiledLHDExpenditures %>%
  filter(year == "2018")

nineteen <- CompiledLHDExpenditures %>%
  filter(year == "2019")
```

```{r}
min(oh_nine$expenditures)

min(ten$expenditures)

min(eleven$expenditures)

min(twelve$expenditures)

min(thirteen$expenditures)

min(fourteen$expenditures)

min(fifteen$expenditures)

min(sixteen$expenditures)

min(seventeen$expenditures)

min(eighteen$expenditures)

min(nineteen$expenditures)
```
```{r}

max(oh_nine$expenditures)

max(ten$expenditures)

max(eleven$expenditures)

max(twelve$expenditures)

max(thirteen$expenditures)

max(fourteen$expenditures)

max(fifteen$expenditures)

max(sixteen$expenditures)

max(seventeen$expenditures)

max(eighteen$expenditures)

max(nineteen$expenditures)
```


```{r}
year <- c('2009', '2010', '2011', '2012', '2013', '2014', '2015', '2016', '2017', '2018', '2019')

high <- c(76972076, 79391383, 79490552, 74369935, 39852467, 38099704, 49579745, 64778296, 69546706, 73079044, 76803875)
low <- c(1166437, 1273596, 1113625, 1041434,  949014, 949693, 892946,  974741, 924449, 900836, 1010820)
data <- data.frame(year, high, low)
data$average <- rowMeans(data[,c("high", "low")])
```

```{r}
linechart <- plot_ly(data, x = year, y = high, type = 'scatter', mode = 'lines',
        line = list(color = 'transparent'),
        showlegend = FALSE, name = 'High')
linechart <- linechart %>% add_trace(y = low, type = 'scatter', mode = 'lines',
            fill = 'tonexty', fillcolor='rgba(0,100,80,0.2)', line = list(color = 'transparent'),
            showlegend = FALSE, name = 'Low') 
linechart <- linechart %>% add_trace(y = data$average, type = 'scatter', mode = 'lines',
            line = list(color='rgb(0,100,80)'),
            name = 'Average')
linechart <- linechart %>% layout(title = "Average, High and Low Expenditures in NC",
         paper_bgcolor='rgb(255,255,255)', plot_bgcolor='rgb(229,229,229)',
         xaxis = list(title = "Years",
                      gridcolor = 'rgb(255,255,255)',
                      showgrid = TRUE,
                      showline = FALSE,
                      showticklabels = TRUE,
                      tickcolor = 'rgb(127,127,127)',
                      ticks = 'outside',
                      zeroline = FALSE),
         yaxis = list(title = "Expenditures",
                      gridcolor = 'rgb(255,255,255)',
                      showgrid = TRUE,
                      showline = FALSE,
                      showticklabels = TRUE,
                      tickcolor = 'rgb(127,127,127)',
                      ticks = 'outside',
                      zeroline = FALSE))
linechart
```



---------------LOLLIPOP CHART---------------------

Dataframe for captia percent change from 2009-2019

```{r}
CapitaPctChange <- CompiledLHDExpenditures %>%
  filter(year %in% c('2009', '2019')) %>%
  pivot_wider(id_cols = -c(expenditures,population), names_from = year, values_from = per_capita_spending) %>%
  mutate(capita_spending_change= (`2019` - `2009`),
    pct_change = 100 * ((`2019` - `2009`)) / `2009`) %>%
  arrange(desc(pct_change))
```


```{r}
ggplot(CapitaPctChange,
       aes(x=pct_change, y=fct_reorder(county_name, pct_change))) +
  geom_segment(
    aes(x=0,
        y=fct_reorder(county_name,pct_change),
        xend= pct_change,
        yend=fct_reorder(county_name, pct_change)),
    color="gray50")+
  geom_point(color="#1d91c0")+
  labs(x="Percent Change in Per Capita Spending", y="County Health Department", 
       title = "Percent Change in Per Capita Spending from 2009-2019",
       caption = "Data Source: County Finance Records") +
  theme_minimal()+
   theme(panel.border = element_blank(),
        panel.grid.minor = element_blank()
   )
```

```{r}
ggsave("lollipop-percent-change.png")
```

