---
title: "Analysis and Charts of LHD expenditure data"
author: "Rachel Crumpler"
output: html_notebook
---

```{r}
#Install packages
library(dplyr)
library(tidyverse)
```

#Importing compiled expenditure data from Google Sheets
```{r}
#install.packages("gsheet")
library(gsheet)
CompiledLHDExpenditures <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Zc10pam92Y1218F90eXn9ri7-a4GQI1vhzING33nNSI/edit#gid=0")
```


#Quick look at column types
```{r}
str(CompiledLHDExpenditures)
```

#Sorting from least expenditures to greatest
```{r}
CompiledLHDExpendituresSorted <- arrange(CompiledLHDExpenditures, desc(expenditures))
```
Mecklenburg, Wake, Guilford have highest expenditures
Jones, Madison and Green have lowest expenditures

#Min, median, mean, and max expenditure values across counties
```{r}
CompiledLHDExpenditures %>% 
  summarise(
  LowestExpenditure= min(expenditures, na.rm=TRUE),
  Median=median(expenditures, na.rm=TRUE),
  Mean=mean(expenditures, na.rm=TRUE),
  HighestExpenditure=max(expenditures, na.rm=TRUE)
  )
```
Highest= Wake 2011
Most recent highest= Mecklenburg 2019 (76,803,875)
Lowest= Jones 2007

#Finding how many counties are greater than mean expenditure value
```{r}
CompiledLHDExpenditures %>%
  filter(expenditures>mean(expenditures, na.rm=TRUE))
```
206 counties greater than mean expenditures

#Min, median, mean, and max per capita spending values across counties
```{r}
CompiledLHDExpenditures %>% 
  summarise(
  LowestCapita= min(per_capita_spending, na.rm=TRUE),
  Median=median(per_capita_spending, na.rm=TRUE),
  Mean=mean(per_capita_spending, na.rm=TRUE),
  HighestCapita=max(per_capita_spending, na.rm=TRUE)
  )
```
Lowest= Stanly County 2001 21.41
Highest= Franklin County 2007

#Sorting from greatest to least of per capita spending
```{r}
PerCapitaSpendingSorted <- arrange(CompiledLHDExpenditures, desc(per_capita_spending))
```
-Franklin County consistently ranks at top in per captia spending with >$250
-Hyde and Dare counties also ranked at top in per capita spending
-Rowan and Stanly counties rank at bottom in per capita spending <$30

#Finding how many counties are greater than mean per capita spending value
```{r}
CompiledLHDExpenditures %>%
  filter(per_capita_spending>mean(per_capita_spending, na.rm=TRUE))
```
275 counties greater than mean per capita spending

#Analyzing expenditure data by year
```{r}
CompiledLHDExpenditures %>%
  group_by(year)%>%
  summarise(
    Mean=mean(expenditures, na.rm=TRUE),
    Median=median(expenditures, na.rm=TRUE),
    SumTotal= sum(expenditures, na.rm=TRUE),
    Total= n()
  )
```
SumTotal= Total expenditures of all counties for that year
Total= Number of counties we have data for

-Median annual expenditures have stayed in the 5,168,918-6,000,000 range from 2007-2020

***Deceiving to look at sum total going up when it is because we have more counties in later years

#Analyzing per capita spending by year
```{r}
CompiledLHDExpenditures %>%
  group_by(year)%>%
  summarise(
    Mean=mean(per_capita_spending, na.rm=TRUE),
    Median=median(per_capita_spending, na.rm=TRUE),
    Min=min(per_capita_spending, na.rm = TRUE),
    Max=max(per_capita_spending, na.rm= TRUE),
    Total= n()
  )
```
Per capita spending grew about $20 from 2000 to 2006. Spending has stayed above at or above $60 since 2006.

-2020 Median= $67
-2019 Median= $68
-2018 Median= $67

Looking at county population sizes
Mecklenburg= largest population
Jones= next to lowest
Hyde= lowest
```{r}
CompiledLHDExpenditures %>%
  arrange(desc(population))
```

----------------PERCENT CHANGE ANALYSIS-------------------

#Looking at expediture percent change per county from 2009 to 2019
```{r}
ExpenditurePctChange <- CompiledLHDExpenditures %>%
  filter(year %in% c('2009', '2019')) %>%
  pivot_wider(id_cols = -c(per_capita_spending,population), names_from = year, values_from = expenditures) %>%
  mutate(expenditure_change= (`2019` - `2009`),
    pct_change = 100 * ((`2019` - `2009`)) / `2009`) %>%
  arrange(desc(pct_change))
```


#Finding average expenditure percent change from 2009 to 2019, min and max
```{r}
ExpenditurePctChange %>%
  summarize(
    AveragePctChange= mean(pct_change, na.rm=TRUE),
    Lowest= min(pct_change, na.rm=TRUE),
    Highest= max(pct_change, na.rm=TRUE)
  )
```
Average percent change from 2009-2019 went up almost 8%
Yadkin and Wake had biggest decrease percent change
Mecklenburg had greatest percent change

#Adding a trend column
```{r}
ExpenditurePctChange <- ExpenditurePctChange %>%

  mutate(trend = case_when(pct_change >0 ~ "up",

                           pct_change < 0 ~ "down",

                           TRUE ~ "no change"

      ))
```

#Seeing how many went up vs. down
```{r}
ExpenditurePctChange %>%
  count(trend)
```
#Filtering counties that only went up 
```{r}
ExpenditurePctChange %>%
  filter(trend== "up")
```
5 counties increased by less than 5%

#Calculation to see when expenditure spending went up, was it by a lot or not?
```{r}
ExpenditurePctChange %>%
  filter(trend== "up") %>%
  summarize(Mean= mean(pct_change),
            Median= median(pct_change),
            Max= max(pct_change),
            Min= min(pct_change)
  )
```

#Filtering counties that went down
```{r}
ExpenditurePctChange %>%
  filter(trend== "down")
```

#Calculation to see when expenditures went down, was it by a lot or not?
```{r}
ExpenditurePctChange %>%
  filter(trend== "down") %>%
  summarize(Mean= mean(pct_change),
            Median= median(pct_change),
            Max= max(pct_change),
            Min= min(pct_change)
  )
```
Went down at a pretty high rate

#Looking at per capita percent change per county from 2009 to 2019
```{r}
CapitaPctChange <- CompiledLHDExpenditures %>%
  filter(year %in% c('2009', '2019')) %>%
  pivot_wider(id_cols = -c(expenditures,population), names_from = year, values_from = per_capita_spending) %>%
  mutate(capita_spending_change= (`2019` - `2009`),
    pct_change = 100 * ((`2019` - `2009`)) / `2009`) %>%
  arrange(desc(pct_change))
```

#Finding per capita percent change from 2009 to 2019, min and max
```{r}
CapitaPctChange %>%
  summarize(
    AveragePctChange= mean(pct_change, na.rm=TRUE),
    Lowest= min(pct_change, na.rm=TRUE),
    Highest= max(pct_change, na.rm=TRUE)
  )
```

#Average Change
```{r}
CapitaPctChange %>%
  summarize(
    AverageChange= mean(capita_spending_change, na.rm=TRUE))
```
Should probably adjust the number for inflation: (inflation rate of 19.2% from 2009-2019) 

Per capita spending on average decreased by $1.65

#Adding trend column
```{r}
CapitaPctChange <- CapitaPctChange %>%

  mutate(trend = case_when(pct_change >0 ~ "up",

                           pct_change < 0 ~ "down",

                           TRUE ~ "no change"

      ))
```

#Seeing how many went up vs. down
```{r}
CapitaPctChange %>%
count(trend)
```
#Filtering counties that only went up 
```{r}
CapitaPctChange %>%
  filter(trend== "up")
```
7 counties per capita spending increased by less than 10%

#Calculation to see when per captia spending went up, was it by a lot or not?
```{r}
CapitaPctChange %>%
  filter(trend== "up") %>%
  summarize(Mean= mean(pct_change),
            Median= median(pct_change),
            Max= max(pct_change),
            Min= min(pct_change)
  )
```

#Filtering counties that went down
```{r}
CapitaPctChange %>%
  filter(trend== "down")
```

#Calculation to see when per captia spending went down, was it by a lot or not?
```{r}
CapitaPctChange %>%
  filter(trend== "down") %>%
  summarize(Mean= mean(pct_change),
            Median= median(pct_change),
            Max= max(pct_change),
            Min= min(pct_change)
  )
```


-------------------- SPECIFIC YEAR ANALYSIS ----------------
#Looking only at 2019 data
```{r}
Data2019 <- CompiledLHDExpenditures %>%
  filter(year == "2019") %>%
  arrange(desc(per_capita_spending))
```
Franklin County highest= $254
Yadkin County lowest= $34

```{r}
Data2019Expenditures <- CompiledLHDExpenditures %>%
  filter(year == "2019") %>%
  arrange(desc(expenditures))
```
Mecklengburg County highest
Jones County lowest


#Closer look at Granville-Vance Health District
*This code can be used to look specifically at other counties of interest
```{r}
CompiledLHDExpenditures %>%
  filter(county_name == "Granville & Vance") %>%
  select(year, expenditures, per_capita_spending) %>%
  mutate(
    Median=median(expenditures),
    Mean=mean(expenditures),
    MedianCapita=median(per_capita_spending),
    MeanCapita=mean(per_capita_spending)
  )
```
*4 mutated columns are not done by year, they represent the value for the county over the years

-Granville-Vance expenditures and per capita spending has gone up gradually over the years


#Closer look at Jones County
```{r}
CompiledLHDExpenditures %>%
  filter(county_name == "Jones") %>%
  select(year, expenditures, per_capita_spending) %>%
  mutate(
    Median=median(expenditures),
    Mean=mean(expenditures),
    MedianCapita=median(per_capita_spending),
    MeanCapita=mean(per_capita_spending)
  )
```










