---
title: "Charts of LHD expenditure data"
author: "Rachel Crumpler"
output: html_notebook
---
#Importing compiled expenditure data from R
```{r}
#install.packages("gsheet")
library(gsheet)
CompiledLHDExpenditures <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Zc10pam92Y1218F90eXn9ri7-a4GQI1vhzING33nNSI/edit#gid=0")
```

#Sorting from least expenditures to greatest
```{r}
CompiledLHDExpendituresSorted <- arrange(CompiledLHDExpenditures, desc(expenditures))
```

#Loading graph packages
```{r}
library("viridis")
#install.packages("hrbrthemes")
library(hrbrthemes)
```

#Making line graph of each individual LHD
```{r}
CompiledLHDExpenditures %>%
  ggplot( aes(x=year, y=expenditures, group=lhd_name, fill=lhd_name)) +
    geom_line() +
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Trends in LHD Expenditures") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    ) +
    facet_wrap(~lhd_name)
```
#Making area chart of each individual LHD
```{r}
CompiledLHDExpenditures %>%
  ggplot( aes(x=year, y=expenditures, group=lhd_name, fill=lhd_name)) +
    geom_area() +
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Trends in LHD Expenditures") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    ) +
    facet_wrap(~lhd_name)
```

#Isolating some of the LHDs for more specific line chart comparison
```{r}
linegraph <- CompiledLHDExpenditures %>% 
  filter(county_name %in% c("Swain","Jones", "Catawba", "Hyde", "Madison", "Mecklenburg"))

# Plot
linegraph %>%
  ggplot( aes(x=year, y=expenditures, group=lhd_name, color=lhd_name)) +
    geom_line() +
    scale_color_viridis(discrete = TRUE) +
    theme(
      legend.position="none",
      plot.title = element_text(size=14)
    ) +
    ggtitle("Trends of LHD expenditures") +
    theme_ipsum()

```


#This graph compares some of the LHD expenditures (can be adjusted for when we find counties we want to put next to each other for closer comparison)
```{r}
linegraph <- CompiledLHDExpenditures %>% 
  filter(county_name %in% c("Swain","Jones", "Catawba", "Hyde", "Madison", "Mecklenburg"))

# Plot
linegraph %>%
  ggplot(aes(x=year, y=expenditures, group=lhd_name, color=lhd_name)) +
    geom_line() +
    scale_color_viridis(discrete = TRUE) +
    theme(
      legend.position="none",
      plot.title = element_text(size=14), 
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    ggtitle("Comparing LHD Expenditures") +
    theme_ipsum()+
    facet_wrap(~lhd_name)
```


#--------THIS JOIN IS NOT CORRECT----------
```{r}
joined_data <- left_join(CompiledLHDExpenditures, NC,
                          by=c("lhd_name" = "lhd_name"))
```

