---
title: "Maps showing Health Department Trends"
author: "Rachel Crumpler"
date: "10/27/2020"
output: html_document
---
Loading packages
```{r}
pacman::p_load(leaflet, glue, dplyr, sf, tmap, tmaptools, tidycensus, ggmap, htmltools,  htmlwidgets, tidyverse)  

pacman::p_load_gh(c("walkerke/tigris", "bhaskarvk/leaflet.extras")) 
```

Downloading shapefile of NC counties
```{r}
options(tigris_use_cache = TRUE)
nc_counties <- counties("NC")
```


Loading our health department data
```{r}
library(gsheet)
CompiledLHDExpenditures <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Zc10pam92Y1218F90eXn9ri7-a4GQI1vhzING33nNSI/edit#gid=0")
```

Renaming column so name is consistent with health data column name so data frames can be joined
```{r}
nc_counties<- dplyr::rename(nc_counties, county_name = NAME) 
```


Joining NC counties shapefile with our health department data
```{r}
library(raster)

# merge on common variable, here called 'key'
county_map <- merge(nc_counties, CompiledLHDExpenditures, by="county_name")

```

#Making a 2009 expenditure map
```{r}
HealthDepartments2009 <- CompiledLHDExpenditures %>%
  filter(year== "2009")
```

```{r}
county_map2009 <- left_join(nc_counties, HealthDepartments2009, by="county_name")
```

#2009 Expenditure Map
*Title isnt displaying
*Need outlines of other counties
```{r}
tmap_mode("view")
tm_shape(county_map2009) +
  tm_polygons("expenditures",
              breaks= c(0, 5000000, 10000000, 15000000, 20000000, 25000000, 30000000,80000000),palette="YlGnBu", title="Fiscal Year 2009 Expenditures")  +
  tm_borders() +
  tm_layout(title = "Health Department Fiscal Year 2009 Spending") +
  tm_legend(legend.position = c("left", "bottom"))
```


#Making a 2019 expenditure map
```{r}
HealthDepartments2019 <- CompiledLHDExpenditures %>%
  filter(year== "2019")
```

```{r}
county_map2019 <- left_join(nc_counties, HealthDepartments2019, by="county_name")
```

```{r}
tmap_mode("view")
tm_shape(county_map2019) +
  tm_polygons("expenditures",
              breaks= c(0, 5000000, 10000000, 15000000, 20000000, 25000000, 30000000,80000000),palette="YlGnBu", title="Fiscal Year 2019 Expenditures")  +
  tm_borders() +
  tm_layout(title = "Health Department Fiscal Year 2019 Spending") +
  tm_legend(legend.position = c("left", "bottom"))
```


#Making 2019 Per Capita Spending Map
*Legend needs to have dollar signs added
*Hover pop ups need more detail
```{r}
tmap_mode("view")
tm_shape(county_map2019) +
  tm_polygons("per_capita_spending", palette="YlGnBu", title="Fiscal Year 2019 Per Capita Spending")  +
  tm_borders() +
  tm_layout(title = "Health Department Fiscal Year 2019 Per Capita Spending") +
  tm_legend(legend.position = c("left", "bottom"))
```


MAKE A MAP SHOWING EXPENDITURE PERCENT CHANGE

Loading in the code of data frame with percent change column
```{r}
ExpenditurePctChange <- CompiledLHDExpenditures %>%
  filter(year %in% c('2009', '2019')) %>%
  pivot_wider(id_cols = -c(per_capita_spending,population), names_from = year, values_from = expenditures) %>%
  mutate(expenditure_change= (`2019` - `2009`),
    pct_change = 100 * ((`2019` - `2009`)) / `2009`) %>%
  arrange(desc(pct_change))
```

Merging shapefile and data
```{r}
# merge on common variable, here called 'key'
pct_map <- left_join(nc_counties, ExpenditurePctChange, by="county_name")
```

Creating map showing percent change
**Need help refining legend position and legend labeling--want them to be percents
```{r}
tm_shape(pct_map) +
  tm_fill("pct_change",title="Percent Change in Health Department Expenditures from 2009-2019",palette="YlGnBu")  +
  tm_borders() +
  tm_layout(title = "Custom Breaks Map", title.position = c("right","bottom"))
```

MAKING A MAP SHOWING PER CAPITA SPENDING PERCENT CHANGE

Loading in the code of data frame with percent change column
```{r}
CapitaPctChange <- CompiledLHDExpenditures %>%
  filter(year %in% c('2009', '2019')) %>%
  pivot_wider(id_cols = -c(expenditures,population), names_from = year, values_from = per_capita_spending) %>%
  mutate(capita_spending_change= (`2019` - `2009`),
    pct_change = 100 * ((`2019` - `2009`)) / `2009`) %>%
  arrange(desc(pct_change))
```

Merging shapefile and data
```{r}
# merge on common variable, here called 'key'
pct_capita_map <- left_join(nc_counties, CapitaPctChange, by="county_name")
```

Creating map showing per capita percent change
**Need help refining legend position and legend labeling--want them to be percents
**Need help refining pop ups ---on hover I want that county's percent change to be displayed
```{r}
tm_shape(pct_capita_map) +
  tm_fill("pct_change",title="Percent Change in Health Department Per Capita Spending from 2009-2019",palette="YlGnBu")  +
  tm_borders() +
  tm_layout(title = "Custom Breaks Map", title.position = c("right","bottom"))
```


MAP SHOWING DOLLAR AMOUNT CHANGE IN PER CAPITA SPENDING FROM 2009-2019
**Need to fix display of consolidated Health Departments
**Need to get hovers showing the dollar change (May show both 2009 and 2019 value as well as the change)
```{r}
tm_shape(pct_capita_map) +
  tm_fill("capita_spending_change",title="Change in Health Department Per Capita Spending from 2009-2019",palette="YlGnBu")  +
  tm_borders()
```

---------------------
Creating map that colored in all counties we have data for — got it to work for the health districts

```{r}
library(gsheet)
MappingData <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1mT0-4-S5YwefLuejmb3GhSN4m5zniLuwzOdAXymOAfk/edit#gid=0")
```

Loading in the code of data frame with percent change column
```{r}
MappingData <- MappingData %>%
  filter(year %in% c('2010', '2018')) %>%
  pivot_wider(id_cols = -c(expenditures,population), names_from = year, values_from = per_capita_spending) %>%
  mutate(capita_spending_change= (`2018` - `2010`),
    pct_change = 100 * ((`2018` - `2010`)) / `2010`) %>%
  arrange(desc(pct_change))
```

Merging shapefile and data
```{r}
# merge on common variable, here called 'key'
new_pct_capita_map <- left_join(nc_counties, MappingData, by="county_name")
```

Creating map showing per capita percent change
**Need help refining legend position and legend labeling--want them to be percents
**Need help refining pop ups ---on hover I want that county's percent change to be displayed
```{r}
tm_shape(new_pct_capita_map) +
  tm_fill("pct_change",title="Percent Change in Health Department Per Capita Spending from 2010-2018",palette="YlGnBu")  +
  tm_borders()
```



